---
title: "Fit the BMORS model for one environment of Maize dataset"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit the BMORS model for one environment of Maize dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This example shows how to evaluate prediction performance with multiple environments and traits in one single environment, using a `BMORS_Env` function. To do this, the MaizeToy dataset is used which is also available in the BMTME package, and it is loaded with the `data()` function as shown below:

```{r}
library(BMTME)
data('MaizeToy')
```

With this command, you load the following 2 objects into the R environment:
  - `genoMaizeToy`: Genomic matrix of the dataset.
  - `phenoMaizeToy`: Phenotypic data from the dataset.

Remember that the first step that we need do, is order the dataset as follows:

```{r}
phenoMaizeToy <- phenoMaizeToy[order(phenoMaizeToy$Env, phenoMaizeToy$Line),]
```

The model to be implemented must be defined and next the matrix design required are obtained with the following code:

```{r}
LG <- cholesky(genoMaizeToy)
ZG <- model.matrix(~0 + as.factor(phenoMaizeToy$Line))
Z.G <- ZG %*% LG
```

Once the matrix design has been generated, the linear predictor is constructed, for this purpose, a new object is created that we call `ETA` that consist of a series of nested lists as shown below,

```{r}
ETA <- list(Gen = list(X = Z.G, model = 'BRR'))
```

Next, we obtain the dataset with the first column with the environments and the others are the phenotypic matrix, this can be done with the following code,

```{r}
dataset <- phenoMaizeToy[, 2:5] #Must Include in the first column the environments
```

Finally, the model will be implemented, and for demonstration purposes only 2000 iterations are used to fit the model. In addition, we will show the resulting predictive performance using the `summary()` function. For this example, we only use the environment `EBU` to check the predictive capacities of the model and use the `covModel = 'BRR'` to specify the Bayes model to be used, by default is Bayes Ridge-Regression, so this parameter could be omitted.

```{r}
pm <- BMORS_Env(dataset, testingEnv = 'EBU', ETA = ETA, covModel = 'BRR', nIter = 2000, 
                  burnIn = 1000, thin = 2, progressBar = TRUE, digits = 3)
summary(pm)
```

To obtain a graph with the Pearson´s correlation or the model we use the `barplot()` function, as shown below,
s

```{r}
barplot(pm)
```

In this figure we can observe that the lowest Pearson´s correlation obtained was observed in the ASI_EBU trait-environment, while the highest Pearson´s correlation obtained in the predictive capacity is observed in the Yield_EBU trait-environment.
If we need recover the predicted values, we could call the results from the object.

```{r}
head(pm$results, 5)
```

Where the Position column, corresponds to the original position in the dataset.